---
title: "Data for Firezer Project"
author: Michael D. Porter
date: June 6, 2023 
format:
  html:
    toc: true
    embed-resources: true
    html-math-method: katex
---

Notes: 

- Updated Nov 15, 2022 (added PTR variables)
- Updated Jan 11, 2023 (using updated data: ECHO, OPO_CTR_CODE)
- Updated Feb 13, 2023 (added more details; no change in data)
- Updated Feb 15, 2023 (changed some variable names)
- Updated Jun 06, 2023 (added donor risk variables, GFR, candidate weight/height)

```{r}
#: Set directories for data and saving
dir_data = "data_in" 
dir_save = "data_out"
```


```{r packages, include=FALSE}
library(knitr)
library(kableExtra)
library(glue)
library(skimr)
library(lubridate)
library(readxl)
library(tidyverse)
library(here)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, 
  out.width = "40%", fig.width=6, fig.height=6,
  message = FALSE, warning = FALSE,
  attr.output='style="max-height: 100px;"', # vertical scroll box
  skimr_include_summary = FALSE
)
options(knitr.kable.NA = '') # set missing table values to blank
options(width = 90)

library(ggplot2)
theme_set(theme_bw())
```

```{r utils, echo=FALSE}
#: make note in html
library(kableExtra)
note <- function(x, col="red") kableExtra::text_spec(x, color=col)

#: round number to specific number of decimal places
digits <- function(x, k=2) formatC(x, digits=k, format="f")

#: print table
library(kableExtra)
print_table <- function(x, digits=3, align="r", position = "left"){
  kbl(x, align = align, digits=digits) %>% 
  row_spec(0, align="c") %>% 
  kable_styling(full_width = FALSE, position = position,
                bootstrap_options = c("striped", "hover","responsive"))
}

#: print scrollable datatable
datatable <- function(x, k=3){
  x %>% 
    mutate_if(is.double, digits, k=k) %>% 
    DT::datatable()
}


#: Calculate time between two date-times (year, month, day, hour, minute, second)
library(lubridate)
calc_dur <- function(DT1, DT2, units = "hour"){
  int = lubridate::interval(DT1, DT2) 
  lubridate::time_length(int, unit = units)  
  #as.numeric(difftime(DT2, DT1, units=units)) # doesn't handle years
}

#: Function to add donor info to measurements
add_donor_info <- function(df, donor_data=donor_data_all,
                           gender=FALSE, age=FALSE){
  DD = donor_data %>% 
    distinct(DONOR_ID, GENDER_DON,AGE_DON, BRAIN_DEATH_DT) # get AGE_DON from PTR
  if(!gender) DD = DD %>% select(-GENDER_DON)
  if(!age) DD = DD %>% select(-AGE_DON)
  df %>%
    left_join(DD, by="DONOR_ID") %>% 
    mutate(hrs_from_DBD = calc_dur(BRAIN_DEATH_DT, DT, "hours")) %>% 
    select(-BRAIN_DEATH_DT)
}

```


# Introduction

Generate data for analysis of OPO donor management. 

## Donor Population

1. Donor age < 18
2. Non DCD Donor
3. Match Run between 2010-2019
    
## Candidate Population

1. All candidates (peds and adults) who were offered a pediatric heart during a qualified match run.  
    
## Data Sets

1. PTR offer data
2. Donor Static data (deceased donor data)
3. DonorNet measurements
4. ECHO measurements


# Data


## Offer Data (PTR)

```{r load-PTR-data}
PTR_all = read_rds(file.path(dir_data, "PTR.rds")) 
# mutate(OPO_CTR_CODE = recode(OPO_CTR_CODE, "10726" = "25581" ))
```

## Donor Attributes (Decesed Donor Data)

```{r load-donor-data}
donor_data_all = read_rds(file.path(dir_data, "donor_data.rds")) %>% 
  semi_join(PTR_all, by="DONOR_ID")
```

## Waitlist Candidate History

```{r load-wl-history}
WL_hist = read_rds(file.path(dir_data, "cand_wl_hist.rds"))
```

## Waitlist Candidate Attributes

```{r load-thoracic}
thoracic = read_rds(file.path(dir_data, "thoracic.rds")) %>% 
  semi_join(PTR_all, by="WL_ID_CODE")
```


## Add additional info to PTR data
```{r update-PTR}
#: Get waitlist status and waitlist organ at time of match run
WL_STATUS = PTR_all %>% select(PTR_ID, WL_ID_CODE, MATCH_SUBMIT_DT) %>% 
  left_join(WL_hist, by="WL_ID_CODE") %>% 
  filter(CHG_DT < MATCH_SUBMIT_DT) %>% 
  group_by(PTR_ID) %>% 
    slice_max(CHG_DT, with_ties = FALSE) %>% 
  ungroup() %>% 
  select(PTR_ID, ORG_CAND = ORG, STATUS_CAND = STATUS)

#: add info to PTR
#   - ORG_CAND, STATUS_CAND
#   - If PTR is missing DOB_DON (and hence AGE_DON), use AGE_DON from donor_data_all
PTR = PTR_all %>% 
  # add ORG_CAND, STATUS_CAND
  left_join(WL_STATUS, by="PTR_ID") %>%
  # add donor age
  left_join(
    donor_data_all %>% select(DONOR_ID, AGE_DON_donor_data = AGE_DON), 
    by = "DONOR_ID"
  ) %>% 
  mutate(
    AGE_DON = coalesce(AGE_DON, AGE_DON_donor_data)
  ) %>% 
  select(-AGE_DON_donor_data, -DON_ID, -CLASS)

rm(WL_hist, WL_STATUS)
```

## Getting Offer Population

There are `r nrow(PTR)` total offers.   

### 1. Match Run between 2010-01-01 and 2019-12-31 (10 years)

```{r PTR1}
study_period = c("2010-01-01", "2019-12-31")
PTR1 = PTR %>% 
  filter(
    as.Date(MATCH_SUBMIT_DT) >= study_period[1], 
    as.Date(MATCH_SUBMIT_DT) <= study_period[2]
  )
```

This leaves a total of `r n_distinct(PTR1$DONOR_ID)` donors, `r n_distinct(PTR1$WL_ID_CODE)` candidates, and `r nrow(PTR1)` offers.

### 2. Pediatric donors (Age at time of match run < 18)

```{r, echo=FALSE}
# Donors with estimated age
tmp = PTR1 %>% mutate(rem = AGE_DON %% 1) %>% filter(rem == 0) %>% 
  select(DONOR_ID)
```

- Note: `r n_distinct(tmp$DONOR_ID)` donors (`r nrow(tmp)` offers) from the PTR records didn't have DOB and thus couldn't calculate `AGE_DON` at time of match run. For these cases, we estimated the age from the `donor_data_all` (deceased donor) records. While this may introduce a few donors that are actually $\geq 18$ at time of match run, there is not expected to be many. 
<!-- - Note: this removes `r PTR1 %>% filter(is.na(AGE_DON)) %>% pull(DONOR_ID) %>% n_distinct` donors, and a total of `r sum(is.na(PTR1$AGE_DON))` offers, with missing DOB (and consequently age) -->

```{r PTR2}
PTR2 = PTR1 %>% filter(AGE_DON < 18)
```

This leaves a total of `r n_distinct(PTR2$DONOR_ID)` donors, `r n_distinct(PTR2$WL_ID_CODE)` candidates, and `r nrow(PTR2)` offers.



### 3. Exclude DCD (donation after cardiac death) donors 

```{r PTR3}
DCD_donors = donor_data_all %>% filter(NON_HRT_DON == "Y") %>% select(DONOR_ID) 
PTR3 = PTR2 %>% anti_join(DCD_donors, by="DONOR_ID")
```

There were `r semi_join(DCD_donors, PTR2, by="DONOR_ID") %>% nrow()` DCD donors, and a total of `r nrow(PTR2) - nrow(PTR3)` offers, that were removed. 

This leaves a total of `r n_distinct(PTR3$DONOR_ID)` donors, `r n_distinct(PTR3$WL_ID_CODE)` candidates, and `r nrow(PTR3)` offers.

### 4. Multiple Match Runs

```{r, echo=FALSE}
multiple_match_runs = PTR3 %>% 
  group_by(DONOR_ID) %>% 
  summarize(
    n_match_runs = n_distinct(MATCH_ID_CODE)
  ) %>% 
  filter(n_match_runs > 1)

multiple_donor_candidate_offers = PTR3 %>% 
  count(DONOR_ID, WL_ID_CODE) %>% filter(n>1)
```

A donor can be offered on multiple match runs. This happens if, for example, test results come back after the initial match run has been initiated. When this occurs, the donor can be offered to the same candidate multiple times since the match run list usually doesn't change much. 

This is an example of a donor who was initially refused by candidate (WL_ID_CODE) 1096112 on the first match run, but accepted only one hour later on the second match run. 
```{r}
PTR %>% filter(DONOR_ID == "433119") %>% 
  select(PTR_ID, DONOR_ID, MATCH_SUBMIT_DT, WL_ID_CODE, PTR_SEQUENCE_NUM, OFFER_ACCEPT) %>%
  arrange(MATCH_SUBMIT_DT, PTR_SEQUENCE_NUM) %>% 
  print_table()
```

This occurred for `r nrow(multiple_match_runs)` of our donors and resulted in `r nrow(multiple_donor_candidate_offers)` donor-candidate pairs that are double (or triple) counted. Options for treating these cases: (a) only use the latest match run, (b) only the last donor-candidate offer,  or (c) count all refusals including multiple same donor-candidate refusals?


```{r, echo=FALSE}
#: keep last donor-candidate offer
# PTR4 = PTR3 %>% 
#   group_by(DONOR_ID, WL_ID_CODE) %>% 
#   slice_max(MATCH_SUBMIT_DT) %>% 
#   ungroup()
```

Currently, we are using option (a) and only keeping offers from the last match run only. 
```{r PTR4}
#: Keep last match run from every donor
PTR4 = PTR3 %>% 
  group_by(DONOR_ID) %>% 
    slice_max(MATCH_SUBMIT_DT, with_ties = TRUE) %>% 
  ungroup()
```

This leaves a total of `r n_distinct(PTR4$DONOR_ID)` donors, `r n_distinct(PTR4$WL_ID_CODE)` candidates, and `r nrow(PTR4)` offers.

## Final Population

- Setting `OFFER_ACCEPT == "Y"` to *Accept* and everything else *Reject*

<!--
- Note: The `OFFER_ACCEPT = Z` is supposed to be provisional acceptance, but it seems that it is the default entry.
    - If candidate doesn't make decision in time, it seems that *Z* is issued. 
    - Check that `INITAL_RESPONSE_DT = FINAL_RESPONSE_TIME`
    - The `PRIME_OPO_REFUSAL_ID` is always missing

```{r, eval=FALSE, echo=FALSE}
PTR %>% 
  filter(OFFER_ACCEPT == "Z") %>% 
  select(INITIAL_RESPONSE_DT, FINAL_RESPONSE_DT, PRIME_OPO_REFUSAL_ID, 
         PTR_SEQUENCE_NUM, MAX_ACCEPT_SEQ) %>% 
  summary
```
-->


```{r PTR-final}
PTR_final = PTR4 %>% 
  mutate(outcome = ifelse(OFFER_ACCEPT == "Y", "Accept", "Reject"))
```

```{r}
PTR_final
```


```{r, eval=FALSE, echo=FALSE}
PTR_final %>% 
  group_by(DONOR_ID, MATCH_ID_CODE) %>% 
  summarize(
    offers = n_distinct(WL_ID_CODE),
    final_outcome = ifelse(any(outcome == "Accept"), "Accept", "Refuse"),
    max_offer = max(PTR_SEQUENCE_NUM)
    # final2 = ifelse(is.na(MAX_ACCEPT_SEQ), "Refuse", "Accept")
  ) %>% ungroup() %>% 
  # count(max_offer - offers2)
   filter(
     max_offer != offers
   )

PTR_final %>% count(WL_ORG)

```


```{r}
PTR_final %>% 
  summarize(
    `offers` = n(),
    `donors` = n_distinct(DONOR_ID), 
    `waitlists` = n_distinct(WL_ID_CODE),
    OPOs = n_distinct(OPO_CTR_CODE)
    #`unique patients` = n_distinct(PT_CODE)
  ) %>% print_table()

PTR_final %>% count(outcome) %>% mutate(p = n/sum(n)) %>% 
  print_table()

```



## Listing Center Volume and Acceptance Rate

- This includes candidates that are still on the waiting list. 
- Counts based on the specified donor/candidate sample. Candidate could have accepted heart from donors not in our sample (e.g., from adult donors). 

```{r}
PTR_final %>% 
  group_by(WL_ID_CODE, LISTING_CTR_CODE) %>% 
    summarize(
      n_offers = n_distinct(DONOR_ID),
      accept = any(outcome == "Accept")
    ) %>% 
  group_by(LISTING_CTR_CODE) %>% 
    summarize(
      n_candidates = n(), 
      n_offers = sum(n_offers),
      n_accepts = sum(accept),
      offers_per_candidate = n_offers / n_candidates,
      offer_accept_rate = n_accepts / n_offers, 
      transplant_rate = n_accepts / n_candidates
    ) %>% 
  ungroup() %>% 
  arrange(-offer_accept_rate) %>% 
  datatable()
```

## OPO Analysis

```{r}

PTR_final %>% 
  group_by(OPO_CTR_CODE) %>% 
    summarize(
      n_donors = n_distinct(DONOR_ID),
      n_offers = n(),    
      n_accepts = sum(outcome == "Accept"),
      offer_accept_rate = n_accepts  / n_offers,
      final_accept_rate = n_accepts / n_donors
    ) %>% 
  ungroup() %>% 
  arrange(-final_accept_rate) %>% 
  datatable()
```

OPO and Listing Center
```{r}
PTR_final %>% 
  count(OPO_CTR_CODE, LISTING_CTR_CODE, sort=TRUE) %>% 
  pivot_wider(names_from = LISTING_CTR_CODE, values_from = n,
              values_fill=0L) %>% 
  datatable()
```

```{r}
PTR_final %>% 
  filter(outcome == "Accept") %>% 
  count(OPO_CTR_CODE, LISTING_CTR_CODE, sort=TRUE) %>% 
  pivot_wider(names_from = LISTING_CTR_CODE, values_from = n,
              values_fill=0L) %>% 
  datatable()
```

# Variable Recoding Functions

Functions for recoding variables that are measured for both donors and candidates.

```{r recoding}

#: Recoding Race/Ethnicity
recode_race <- function(ETHCAT){
  recode(ETHCAT, 
          `1` = "White", 
          `2` = "Black", 
          `4` = "Hispanic", 
          #`5` = "Asian", 
          #`6` = "Amer Ind/Alaska Native", 
          #`7` = "Native Hawaiian/other Pacific Islander",
          #`9` = "Multiracial",
          .default = "Other", .missing = "Unknown"
          )  
}

#: Recoding Blood Type
recode_ABO <- function(ABO){
  recode(ABO, A1 = "A", A2 = "A", A1B = "AB", A2B = "AB", 
         .missing = "Unknown")  
}

#: Recode Y = Yes, N = No, U = Unknown
recode_YNU <- function(YNU){
  recode(YNU, N = "No", Y = "Yes", U = "Unknown", .missing = "Unknown")
}

#: Recode P = Yes, N = No, anything else = Unknown
recode_PNO <- function(PNO){
  recode(PNO, N = "No", P = "Yes", .default = "Unknown", .missing = "Unknown")
}
# P = Positive
# N = Negative
# C = Cannot Disclose
# I = Indeterminate
# ND = Not Done
# PD = Pending
# U = Unknown	
# OTHER = Unknown

```

# Donor Attributes

- Gender `GENDER_DON`
- Race `RACE_DON`
- Weight (`WEIGHT_DON_KG`), Height (`HEIGHT_DON_CM`)
- Blood type `ABO_DON`
- Cause of Death `COD_DON`
- Donor history of chest trauma `CHEST_TRAUMA`
- CPR
    a. History of CPR before and/ or after declaration of brain death (I believe that’s a category)
    b. Duration of CPR
    c. Cardiac arrest total downtime (variable within deceased donor data)
- Donor Risk Factors (drug use)



```{r donor_data}
donor_data = donor_data_all %>% 
  semi_join(PTR_final, by="DONOR_ID") %>% 
  transmute(
    #: Donor ID
    DONOR_ID,
    #: Gender {M,F}
    GENDER_DON,   
    #: Ethnicity
    RACE_DON = recode_race(ETHCAT_DON),
    #: Weight in kg
    WEIGHT_DON_KG = WGT_KG_DON_CALC, 
    #: Height in cm
    HEIGHT_DON_CM = HGT_CM_DON_CALC,
    #: Blood Type
    ABO_DON = recode_ABO(ABO_DON),
    #: Did donor have CPR?
    CPR_ADMIN = recode_YNU(CPR_ADMIN),
    #: Duration of CPR (in mins) for donors with CPR
    CPR_DURATION = ifelse(CPR_ADMIN == "No", 0, CPR_ADMIN_DURATION),
    #: Cardiac Arrest Downtime? (Y=Yes; N=No; U=Unknown) 
    CARDARREST = recode_YNU(CARDARREST_DOWNTM),
    #: Cardiac Arrest Downtime Duration (in Minutes)
    CARDARREST_DURATION = ifelse(CARDARREST == "No", 0, CARDARREST_DOWNTM_DURATION),
    #: Cardiac arrest after brain death:
    CARDARREST_POSTNEURO = recode_YNU(CARDARREST_POSTNEURO_DON),
    #: Time of Brain Death
    BRAIN_DEATH_DT,
    #: Donor Cause of Death
    COD_DON,
    #: Donor Circumstance of Death
    DEATH_CIRCUM_DON,
    #: Donor Mechanism of Death
    DEATH_MECH_DON,
    #: Donor history of Chest Trauma
    CHEST_TRAUMA = recode_YNU(CHEST_TRAUMA),

    #: Increased Risk Factors
    HIST_COCAINE_DON = recode_YNU(HIST_COCAINE_DON), 
    CURRENT_COCAINE_DON = recode_YNU(CONTIN_COCAINE_DON),
    HIST_IV_DRUG_DON = recode_YNU(HIST_IV_DRUGUSE), 
    HIST_OTHER_DRUG_DON = recode_YNU(HIST_OTH_DRUG_DON), 
    CURRENT_OTHER_DRUG_DON = recode_YNU(CONTIN_OTH_DRUG_DON),
    HIST_CIG_DON = recode_YNU(HIST_CIG_DON), 
    CURRENT_CIG_DON = recode_YNU(CONTIN_CIG_DON), # Cigarette Use > 20 Pack Years Ever
    # HIST_CAD_DON = recode_YNU(HIST_CAD), # (n=3) History of Coronary Artery Disease
    HIST_DRUG_DON = case_when(
      HIST_COCAINE_DON  == "Y" ~ "Y", 
      HIST_IV_DRUGUSE   == "Y" ~ "Y",
      HIST_OTH_DRUG_DON == "Y" ~ "Y",
      (HIST_COCAINE_DON == "N" & HIST_IV_DRUGUSE == "N" & 
         HIST_OTH_DRUG_DON =="N") ~ "N",
      .default = "U"
    ) %>% recode_YNU(), 
    CURRENT_DRUG_DON = case_when(
      CONTIN_COCAINE_DON  == "Y" ~ "Y", 
      CONTIN_OTH_DRUG_DON == "Y" ~ "Y",
      (CONTIN_COCAINE_DON == "N" & CONTIN_OTH_DRUG_DON == "N") ~ "N",
      .default = "U"
    ) %>% recode_YNU(), 
    
    #: Infectious Disease
    RISK_CDC_DON = recode_YNU(CDC_RISK_HIV_DON), # Per PHS, Does the Donor Have Risk Factors for Blood-Borne Disease Transmission?
    HBV_CORE_DON = recode_PNO(HBV_CORE_DON), # Hep B Core
    HBV_NAT_DON = recode_PNO(HBV_DNA),       # Hep B NAT
    HBV_SUR_ANTIGEN_DON = recode_PNO(HBV_SUR_ANTIGEN_DON), # Hep B Surface Antigen
    HBV_Ab_DON = recode_PNO(HBSAB_DON),      # Hep B Antibodies    
    HCV_NAT_DON = recode_PNO(HCV_NAT),       # Hep C NAT
    HCV_Ab_DON = recode_PNO(HEP_C_ANTI_DON), # Hep C Antibodies
    CMV_DON = recode_PNO(CMV_DON),           # Cytomegalovirus (CMV) antibody 
    # TOXO_IGG_DON
    # EBV_IGG_CAD_DON
    # EBV_IGM_CAD_DON
    # HTLV_DON,
    # HTLV_NAT 
    #: HIV (Note: only 4 total positives in all donor data)
    # HIV_ANTIGEN_DON = recode_PNO(HIV_ANTIGEN) # HIV Ag/Ab Combo Assay
    # HIV_SER_DON = recode_PNO(HIV_DON)         # HIV Serology Results
    # HIV_NAT_DON = recode_PNO(HIV_NAT_DON)     # HIV NAT

)
```


```{r}
donor_data %>% mutate_if(is.character, factor) %>% 
  skimr::skim()
```


```{r donor_data-check, echo=FALSE}
if(n_distinct(donor_data$DONOR_ID) != nrow(donor_data)){
  stop("donor_data contained duplicate entries. Need to fix!")
}
```



## Donor Cause of Death 

- See the file `COD-text.csv` for what we code all free text descriptions

```{r donor-COD}
donor_data %>% count(COD_DON) %>% print_table()
```

```{r}
## Check COD classification
donor_data_all %>%
  semi_join(PTR_final, by="DONOR_ID") %>%
  filter(!is.na(COD_OSTXT_DON)) %>%
  count(COD_OSTXT_DON, COD_DON, sort=TRUE) %>% 
  print_table() %>% scroll_box(height = "400px")
```

 

# DonorNet Measurements

## VIS and Thyroid Repletion Therapy (T4)

- `hrs_from_DBD` is hours since declaration of brain death
- `VIS` is Vasoactive-Inotropic Score
- `Thyroid` {0, 1} for patient receiving Thyroid (Levothyroxine, T4, T3) (Thyroid Repletion Therapy)
- `Dobutamine` {0, 1} for patient receiving Dobutamine
- `Dopamine` {0, 1} for patient receiving Dopamine
- `Epinephrine` {0, 1} for patient receiving Epinephrine
- `Milrinone` {0, 1} for patient receiving Milrinone
- `Norepinephrine` {0, 1} for patient receiving Norepinephrine
- `Vasopressin` {0, 1} for patient receiving Vasopressin
- `INO_count` number of simultaneous inotropes being administered to donor

- Note: Missing VIS replaced with 0. 

```{r Inotropes}
INOTROPES = read_rds(file.path(dir_data, "INOTROPES.rds")) %>% 
  semi_join(PTR_final, by="DONOR_ID") %>% 
  add_donor_info() %>% 
  transmute(
    DONOR_ID, DT, hrs_from_DBD,
    VIS, 
    Thyroid, 
    Dobutamine  = 1L*(Dobutamine > 0),
    Dopamine    = 1L*(Dopamine > 0),
    Epinephrine = 1L*(Epinephrine > 0),
    Milrinone   = 1L*(Milrinone > 0),
    Norepinephrine = 1L*(Norepinephrine > 0),
    Vasopressin = 1L*(Vasopressin > 0),
    INO_count = 0 + (Dobutamine > 0) + (Dopamine > 0) + (Epinephrine > 0) + 
      (Milrinone > 0) + (Norepinephrine > 0) + (Vasopressin > 0)
  )

```


```{r}
INOTROPES
```


```{r}
INOTROPES %>% count(DONOR_ID) %>% 
  complete(DONOR_ID = unique(PTR_final$DONOR_ID), fill = list(n = 0L)) %>% 
  ggplot(aes(n)) + geom_histogram(boundary = 0, binwidth=5) + 
  # scale_y_log10() + 
  labs(x = "number of inotrope measurements per donor")
```

## Pulse

- Using age-adjusted thresholds specified below. `r note('Unsure of source.')`

```{r pulse-ranges}
#: Set pulse thresholds by age
pulse_range = tribble(
~ age_group,  ~age_cut, ~thres,
'< 1',    0, 160,
'1-2',    2, 150,
'3-5',    5, 120, 
'6-11',  11, 110,
'12+',  Inf, 100
)

pulse_range %>% 
  select(`Age Group` = age_group, threshold=thres) %>%  
  print_table()
```

- Using `AVG_PULSE_RANGE_{START, END}` as pulse measurement
- `PULSE_MIN`, `PULSE_MAX` Minimum and Maximum pulse
- `PULSE_high` 1 if pulse exceeds age-adjusted threshold, 0 otherwise
- `hrs_from_DBD` is hours since declaration of brain death

```{r PULSE_donor}
PULSE = read_rds(file.path(dir_data, "PULSE.rds")) %>% 
  semi_join(PTR_final, by="DONOR_ID") %>% 
  add_donor_info(age=TRUE) %>% 
  mutate(age_group = cut(floor(AGE_DON), 
                         breaks = c(-Inf, pulse_range$age_cut),
                         labels = pulse_range$age_group,  
                         right = TRUE)) %>% 
  left_join(pulse_range, by="age_group") %>% 
  select(-age_group, -age_cut) %>% 
  mutate(PULSE_high = 1L*(PULSE_MAX >= thres))
```

```{r}
PULSE
```

## Central Venous Pressure (CVP)

- Data from: `CVP_INT_RANGE_{START, END}`
- `CVP_MIN`, `CVP_MAX`
- `hrs_from_DBD` is hours since declaration of brain death

```{r CVP_donor}
CVP = read_rds(file.path(dir_data, "CVP.rds")) %>% 
  semi_join(PTR_final, by="DONOR_ID") %>% 
  add_donor_info()
```

```{r}
CVP
```

## Blood Pressure

- Original Request: Proportion of time systolic blood pressure is greater than 2 S.D. greater than the upper limits of normal

- But we don't have S.D., so using percentiles to determine low blood pressure (< 5th percentile) and high blood pressure (> 95th percentile)

- We are using Average Systolic Blood Pressure: The average BP over some time period.

- `r note('I don\'t have the SD units, so using time above the ranges specified in the variable_ranges spreadsheet: see below')`


- Reference for thresholds:
    - Electronic Supplementary Material
    - Risk Factors and Inpatient Outcomes Associated with Acute Kidney Injury at Pediatric Severe Sepsis Presentation 
    - Julie C. Fitzgerald, MD, PhD1,2; Michelle E. Ross, PhD3; Neal J. Thomas, MD, MSc4; Scott L. Weiss, MD MSCE1, Fran Balamuth, MD, PhD, MSCE5,6, and Amanda Hyre Anderson, PhD3
    - Using the 95th percentile as high threshold and 5th percentile as low percentile
    - Source: <https://sites.google.com/a/channing.harvard.edu/bernardrosner/pediatric-blood-press/boys-percentiles-of-blood-perssure-by-percentile-of-height>
    
    
```{r, eval=FALSE, echo=FALSE}
# url = "https://sites.google.com/a/channing.harvard.edu/bernardrosner/pediatric-blood-press/boys-percentiles-of-blood-perssure-by-percentile-of-height/1-year-old-1"
# library(rvest)
# x = read_html(url)
# x %>% html_element("body > table") #?? not working. iframe?
```
    

```{r bp-ranges}
#: Set blood pressure thresholds by age and gender

library(readxl)
bp_range = bind_rows(
  "F" = read_excel(file.path(dir_data, "ranges/Systolic_BP.xls"), 
                   sheet = "girls"),
  "M" = read_excel(file.path(dir_data, "ranges/Systolic_BP.xls"), 
                   sheet = "boys"), 
  .id = "Gender"
) %>% 
  select(GENDER_DON = "Gender", AGE_DON = "Age", low = "5th", high = "95th") %>%
  # add Age = 0 and Age = 18
  bind_rows(
    filter(., AGE_DON == 1) %>% mutate(AGE_DON = 0),
    filter(., AGE_DON == 17) %>% mutate(AGE_DON = 18)
  ) %>% 
  arrange(GENDER_DON, AGE_DON)
```

```{r}
bp_range %>% print_table() %>%
  scroll_box(height = "400px")
```


- Using `AVG_BP_SYST` as measurement
- `BP_high` is 1 if measurement exceeds threshold
- `BP_low` is 1 if measurement is too low
- `hrs_from_DBD` is hours since declaration of brain death

```{r BP_donor}

BP = read_rds(file.path(dir_data, "BP.rds")) %>% 
  semi_join(PTR_final, by="DONOR_ID") %>% 
  add_donor_info(donor_data_all, gender=TRUE, age=TRUE) %>% 
  mutate(AGE_DON = floor(AGE_DON)) %>% # ranges use integers
  left_join(bp_range, by=c("AGE_DON", "GENDER_DON")) %>% 
  mutate(
    BP_high = 1L*(AVG_BP_SYST > high),
    BP_low  = 1L*(AVG_BP_SYST < low)
  ) 
```

```{r}
BP
```


```{r}
BP %>% 
  summarize(
    n_missing = sum(is.na(AVG_BP_SYST)),
    percent_high = mean(BP_high, na.rm=TRUE),
    percent_low = mean(BP_low, na.rm=TRUE)
  )
```
Not sure if this has something to do with donor management, but about 5% of BP are at the low threshold (5th percentile), but 37% exceed the high threshold (95th percentile). This indicates that donors are kept at a higher blood pressure than general pediatric population.




## Labs


- Labs 
    a. Hgb (Hemogoblin) \[CBC\]
    b. Hct (Hematocrit) \[CBC\]
    c. Lowest eGFR (Estimated Glomerular Filtration Rate) \[LABS\]
    d. Highest PELD or MELX-XI score \[LABS\]
    e. Lowest pH \[ABGS:ABG_PH\]
    f. Lowest paO2 (Partial Pressure of Oxygen) \[ABGS\]
    g. PEEP \[ABGS\]
    h. Highest Creatinine \[LABS\]
    i. Peak and final Troponin I value (it is typically elevated after CPR)

- TODO: Check records for incorrect dates

```{r LABS_donor}
LABS = read_rds(file.path(dir_data, "LABS.rds")) %>% 
  semi_join(PTR_final, by="DONOR_ID") %>% 
  add_donor_info()
ABGS = read_rds(file.path(dir_data, "ABGS.rds")) %>% 
  semi_join(PTR_final, by="DONOR_ID") %>% 
  add_donor_info()
CBC = read_rds(file.path(dir_data, "CBC.rds")) %>% 
  semi_join(PTR_final, by="DONOR_ID") %>% 
  add_donor_info()

LAB_data = bind_rows(LABS, ABGS, CBC)
```

```{r}
LAB_data
```


```{r}
#!!! BAD RECORDS
LAB_data %>% 
  filter(DT <= "1999-01-01") %>% 
  distinct(DONOR_ID, DT) %>% print_table()
```





# ECHOS


## Donor Echos (ECHO) 

```{r load-echo-data}
echo = read_rds(file.path(dir_data, "echo.rds")) %>% 
  semi_join(PTR_final, by="DONOR_ID") %>% 
  select(-DON_ID)
```

```{r}
echo
```




### Quality Checks

- Note: some records were opened, but there was no echo data. 

```{r}
echo_missing = PTR_final %>% 
  anti_join(echo, by="DONOR_ID") %>% 
  summarize(
    n_donors = n_distinct(DONOR_ID), 
    p_donors = n_donors / n_distinct(PTR_final$DONOR_ID),
    n_offers = length(DONOR_ID), 
    p_offers = n_offers / nrow(PTR_final),
    n_candidates = n_distinct(WL_ID_CODE)
  )

#: missing dates
echo_missing_dates = echo %>% 
  semi_join(PTR_final, by="DONOR_ID") %>% 
  summarize(missing_dates = sum(is.na(DT)))
```


- There are `r pull(echo_missing_dates)` echos for our donor population that are missing dates/times. 
    - Note: these all correspond to records that should have been deleted. Safe to ignore.  

- There are `r echo_missing$n_donors` (`r round(100*echo_missing$p_donors)`%) *donors with no ECHOs*. 


- There are `r echo_missing$n_offers` (`r round(100*echo_missing$p_offers)`%) *offers with no ECHOs*. 




### Classify ECHO
```{r echo_data}
#: classify echos as normal or abnormal
echo_data = echo %>% 
  filter(!is.na(DONOR_ID), !is.na(DT)) %>% 
  arrange(DONOR_ID, DT) %>% 
  #: extract integer codes
  mutate(
    across(
      c(mitral_insuff, aortic_insuff, tricuspid_insuff, pulmonary_insuff, 
        global_left_vent_dysf, global_right_vent_dysf, 
        focal_leftvent_free_wall_dysf, focal_intervent_septal_dysf), 
      ~ str_extract(., "\\d") %>% as.integer
    )) %>% 
  #: LVSWMA  
  mutate(
    LVSWMA = case_when(
      focal_leftvent_free_wall_dysf > 0 ~ 1L, 
      focal_intervent_septal_dysf   > 0 ~ 1L, 
      TRUE ~ 0L  # both are 0 or missing (or one is 0, other is missing)
  )) %>% 
  #: Objective Measurements
  mutate(
    ECHO_OBJECTIVE = case_when(
      # first check biplane_eject_frac
      biplane_eject_frac >= 55 ~ 0L, 
      biplane_eject_frac <  55 ~ 1L, 
      # else biplane_eject_frac is missing
      four_chamber_eject_frac >= 55 ~ 0L, 
      four_chamber_eject_frac <  55 ~ 1L, 
      # else biplane_eject_frac & four_chamber_eject_frac are missing
      shortening_frac >= 28 ~ 0L,
      shortening_frac <  28 ~ 1L,
      # if all are missing, then leave as NA
    )
  ) %>% 
  #: Objective Scale Measurements
  mutate(
    ECHO_OBJECTIVE_SCALE = case_when(
      # first check biplane_eject_frac
      biplane_eject_frac >= 55                  ~ 0L, # [55, 100]
      between(biplane_eject_frac, 45, 55)       ~ 1L, # [45, 55)
      between(biplane_eject_frac, 35, 45)       ~ 2L, # [35, 45)
      biplane_eject_frac < 35                   ~ 3L, # [0, 35)
      # else biplane_eject_frac is missing
      four_chamber_eject_frac >= 55             ~ 0L, # [55, 100]
      between(four_chamber_eject_frac, 45, 55)  ~ 1L, # [45, 55)
      between(four_chamber_eject_frac, 35, 45)  ~ 2L, # [35, 45)
      four_chamber_eject_frac < 35              ~ 3L, # [0, 35)  
      # else biplane_eject_frac & four_chamber_eject_frac are missing
      shortening_frac >= 28             ~ 0L, # [28, 100]
      between(shortening_frac, 23, 28)  ~ 1L, # [23, 28)
      between(shortening_frac, 18, 23)  ~ 2L, # [18, 23)
      shortening_frac < 18              ~ 3L, # [0, 18)        
      # if all are missing, then leave as NA
    )
  ) %>% 
  #: VALVE_FXN
  mutate( 
    VALVE_FXN = case_when(
      aortic_insuff >= 2    ~ 1L,
      mitral_insuff >= 2    ~ 1L,
      tricuspid_insuff >= 2 ~ 1L,
      pulmonary_insuff >= 2 ~ 1L,
      TRUE ~ 0L)   # anything else, including all missing (NA)
  ) %>%  
  #: Qualitative Echo
  mutate(
    ECHO_QUAL = case_when(
      global_left_vent_dysf == 0 & qual_eject_frac <  55 ~ 0L, 
      global_left_vent_dysf == 1 & qual_eject_frac >= 55 ~ 0L, 
      global_left_vent_dysf >  0 ~ 1L,
      qual_eject_frac < 55 ~ 1L, 
      is.na(qual_eject_frac) & is.na(global_left_vent_dysf) ~ NA_integer_,
      TRUE ~ 0L
    )
  ) %>% 
  mutate(ABNL_ECHO_CUM = case_when(
    # if any are bad, then bad
    ECHO_OBJECTIVE == 1         ~ 1L,
    ECHO_QUAL == 1              ~ 1L,
    LVSWMA == 1                 ~ 1L,
    VALVE_FXN == 1              ~ 1L,
    global_left_vent_dysf > 0   ~ 1L, 
    global_right_vent_dysf > 0  ~ 1L, 
    # if all missing, then NA
    is.na(ECHO_OBJECTIVE) & is.na(ECHO_QUAL) & 
      is.na(LVSWMA) & is.na(VALVE_FXN) &
      is.na(global_left_vent_dysf) & is.na(global_right_vent_dysf) ~ NA_integer_, 
    # otherwise, its good
    TRUE ~ 0L)
  ) %>% 
  select(DONOR_ID, DT, LVSWMA, ECHO_OBJECTIVE, ECHO_OBJECTIVE_SCALE, 
         VALVE_FXN, ECHO_QUAL, ABNL_ECHO_CUM) %>% 
  add_donor_info()
```

```{r}
echo_data
```



```{r}
echo_data %>% 
  select(-DONOR_ID, -DT) %>% 
  skimr::skim()
```
Number of Echos per donor
```{r}
echo_data %>% 
  count(DONOR_ID) %>% 
  complete(
    DONOR_ID = unique(PTR_final$DONOR_ID),
    fill = list(n = 0L)       
  ) %>% 
  count("echos_per_donor" = n, name="n") %>% mutate(p=n/sum(n)) %>% 
  print_table()
```

# Data for analysis

- Keep measurements after declaration of brain death (`hrs_from_DBD` > 0)

```{r}
save_it <- function(x, nm){
  nm = file.path(dir_save, nm)
  write_csv(x, nm)
}
```


```{r}
thoracic %>% 
  transmute(
    WL_ID_CODE,
    WEIGHT_CAND_KG = coalesce(WGT_KG_CALC, WGT_KG_TCR), 
    HEIGHT_CAND_CM = coalesce(HGT_CM_CALC, HGT_CM_TCR),
  ) %>% 
  save_it("candidate_data.csv")
```


```{r}

PTR_all %>% save_it("PTR_all.csv")

#: PTR data
PTR_final %>% 
  transmute(PTR_ID,
    # donor info
    DONOR_ID, AGE_DON, OPO_CTR_CODE, 
    # candidate info
    WL_ID_CODE, AGE_PT, LISTING_CTR_CODE, STATUS_CAND, 
    # refusal codes
    REFUSAL_CD = PRIME_OPO_REFUSAL_ID %>% coalesce(TXC_REFUSAL_CD) %>% 
      coalesce(INITIAL_REFUSAL_CD),
    REFUSAL_CD = ifelse(outcome == "Accept", NA, REFUSAL_CD),
    # accept/reject outcome
    outcome, MAX_ACCEPT_SEQ,
    # offer info
    DISTANCE, MATCH_ID_CODE, MATCH_SUBMIT_DT, PTR_SEQUENCE_NUM,
  ) %>% 
  save_it("PTR_final.csv")


#: donor_data
donor_data %>% save_it("donor_data.csv")
  
#: inotropes
INOTROPES %>% 
  filter(hrs_from_DBD > 0) %>% 
  group_by(DONOR_ID) %>% 
    summarize(
      n = n(),
      VIS_peak = max(VIS),
      INO_count_max = max(INO_count),
      Thyroid_ever = max(Thyroid)
    ) %>% 
  save_it("Inotropes.csv")
  
#: blood-pressure
BP %>% 
  filter(hrs_from_DBD > 0) %>% 
  group_by(DONOR_ID) %>% 
    summarize(
      n = n(),
      BP_high_prop = mean(BP_high),
      BP_low_prop = mean(BP_low)
    ) %>% 
  save_it("blood-pressure.csv")

#: troponin
LAB_data %>% 
  filter(hrs_from_DBD > 0, variable == "TROPONINI") %>% 
  group_by(DONOR_ID) %>%
    summarize(
      n = n(),
      TROPONIN_peak = max(value), 
      TROPONIN_avg = mean(value),
      TROPONIN_med = median(value)
    ) %>% 
  save_it("troponin.csv")


#: GFR
LAB_data %>% 
  filter(hrs_from_DBD > 0, variable == "GFR") %>% 
  group_by(DONOR_ID) %>%
    summarize(
      n = n(),
      GFR_min = min(value), 
      GFR_avg = mean(value),
      GFR_med = median(value)
    ) %>%
  save_it("gfr.csv")

#: echo
echo_data %>% 
  filter(hrs_from_DBD > 0) %>% 
  group_by(DONOR_ID) %>%
    arrange(DT) %>% 
    summarize(
      n = n(),
      echo_final_abnormal = tail(ABNL_ECHO_CUM > 0, n=1),
      echo_any_abnormal = any(ABNL_ECHO_CUM > 0),
      echo_any_normal = any(ABNL_ECHO_CUM < 1),
      echo_improves = any(lead(ABNL_ECHO_CUM) > ABNL_ECHO_CUM, na.rm=TRUE),
      echo_worsens = any(lead(ABNL_ECHO_CUM) < ABNL_ECHO_CUM, na.rm=TRUE),
    ) %>% 
  save_it("echo.csv")
```


```{r}
#: listing center acceptance rates
PTR_final %>% 
  group_by(WL_ID_CODE, LISTING_CTR_CODE) %>% 
    summarize(
      n_offers = n_distinct(DONOR_ID),
      accept = any(outcome == "Accept")
    ) %>% 
  group_by(LISTING_CTR_CODE) %>% 
    summarize(
      n_candidates = n(), 
      n_offers = sum(n_offers),
      n_accepts = sum(accept),
      offers_per_candidate = n_offers / n_candidates,
      offer_accept_rate = n_accepts / n_offers, 
      transplant_rate = n_accepts / n_candidates
    ) %>% 
  save_it("listing-center.csv")


#: OPO center acceptance rates
PTR_final %>% 
  group_by(OPO_CTR_CODE) %>% 
  summarize(
    n_donors = n_distinct(DONOR_ID),
    n_offers = n(),    
    n_accepts = sum(outcome == "Accept"),
    offer_accept_rate = n_accepts  / n_offers,
    final_accept_rate = n_accepts / n_donors
  )  %>% 
  save_it("opo.csv")
  

#: OPO x Listing center 
PTR_final %>% 
  group_by(OPO_CTR_CODE, LISTING_CTR_CODE) %>% 
    summarize(
      n_offers = n(), 
      n_accepts = sum(outcome == "Accept")
    ) %>% 
  save_it("opo-listing.csv")

```



